Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system of plausible mechanism

### 2024-10-17 index.qmd
section:introduction; process:write

Over the past decade, numerous studies have documented the effectiveness of the *comparative judgment* (CJ) method [@Thurstone_1927; @Pollitt_2012a] for assessing competencies and traits. These studies have evaluated CJ from two main perspectives: its ability to produce reliable and valid trait scores, and its practical applicability. In terms of reliability and validity, research has shown that CJ can generate precise and consistent scores that accurately represent the traits being measured. Notable contributions in this research area include studies by @Pollitt_2012b, @Whitehouse_2012, @vanDaal_et_al_2016, @Lesterhuis_2018, @vanDaal_et_al_2019, @Bramley_et_al_2019, @Verhavert_et_al_2019, @Crompvoets_et_al_2022, and @Bouwer_et_al_2023. Regarding practical applicability, several studies have highlighted the method's versatility in both educational and non-educational contexts, presenting it as an efficient and effective alternative for measurement and evaluation. Key examples in this research area include the works of @Jones_2015, @Bartholomew_et_al_2018, @Jones_et_al_2019, @Marshall_et_al_2020, @Bartholomew_et_al_2020, and @Boonen_et_al_2020.

@@

Nevertheless, despite the growing number of CJ studies, the research approaches employed in the literature have been unsystematic and non-integrated, leading to the oversight of several critical issues related to the method. These issues include concerns about the measurement model responsible for generating the CJ scores, the structural component used for further data analysis and hypothesis testing, and challenges related to the design of CJ experiments. For instance, a notable concern regarding the measurement model is the prevalent reliance on the assumptions of Case 5 from Thurstone's law of comparative judgment [-@Thurstone_1927]. Although Case 5 was originally articulated to produce a "rough measurement" or "rather coarse scaling" of traits [@Thurstone_1927, p. 268-269], its assumptions have become predominant in the literature due to its implementation through the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959]. This leaves issues such as the presence of judge' biases hinted by @Bramley_2008 and @Kelly_et_al_2022, and evidenced by @Pollitt_et_al_2003, @Arlett_2003, and @Guthrie_2003

<!-- a practice that may not be justifiable. -->

<!-- One issue related to the structural component, is the disconnection existing in the literature between the measurement model and the structural component used for hypothesis testing,  -->

@@

<!-- possibility of bias is hinted in @Bramley_2008, and later mentioned again in @Kelly_et_al_2022, and some evidence is found in @Pollitt_et_al_2003, @Arlett_2003, @Guthrie_2003 and @Gijsen_et_al_2021 -->

>>>>

Over the past decade, numerous studies have documented the effectiveness of the *comparative judgment* (CJ) method [@Thurstone_1927] for assessing competencies and traits. These studies have evaluated CJ from two main perspectives: its ability to produce reliable and valid trait scores, and its practical applicability. In terms of reliability and validity, research has shown that CJ can generate precise and consistent scores that accurately represent the traits being measured. Notable contributions in this research area include studies by Pollit [-@Pollitt_2012a; -@Pollitt_2012b], @Whitehouse_2012, @vanDaal_et_al_2016, @Lesterhuis_2018, @Bramley_et_al_2019, @Verhavert_et_al_2019, @Crompvoets_et_al_2022, and @Bouwer_et_al_2023. Regarding practical applicability, several studies have highlighted the method's versatility in both educational and non-educational contexts, presenting it as an efficient and effective alternative for measurement and evaluation. Key examples in this research area include the works of @Jones_2015, @Bartholomew_et_al_2018, @Jones_et_al_2019, @Marshall_et_al_2020, @Bartholomew_et_al_2020, and @Boonen_et_al_2020.

@@

Despite the growing number of CJ studies, the unsystematic and fragmented research approaches employed in the literature have overlooked several critical issues concerning CJ. These issues can be categorized into concerns about the method's structural model, measurement model, and experimental design.

A key issue in the first category is the apparent disconnect between the method's structural and measurement models. A common practice in CJ literature involves performing data analysis and hypothesis testing on scores previously estimated using the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959]. These studies use the scores or their transformations to identify 'misfit' judges and stimuli [@Pollitt_2012b; @vanDaal_et_al_2017], detect 'bias' in judges' ratings [@Pollitt_et_al_2003; @Pollitt_2012b], or test various hypothesis [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021]. However, since CJ scores are parameter estimates with inherent uncertainty, the statistical literature suggests that separating the analysis from this uncertainty may artificially inflate the precision and power of the results. This, in turn, could increase the risk of committing a type I error, where a null hypothesis is wrongly rejected [@McElreath_2020].

<!-- Moreover, a notable concern in the second category is the measurement model's reliance on the assumptions of Case 5 from Thurstone's law of comparative judgment [-@Thurstone_1927]. As noted, it is common practice in the literature to use the BTL model as a means of generating CJ scores, serving as the measurement model. -->

<!-- Since the study by @Bramley_2008, it has been clear that the BTL model represents a statistical articulation of Case 5. -->

<!-- Although Case 5 was originally articulated to produce a "rather coarse scaling" of traits [@Thurstone_1927, p. 269], its assumptions have become predominant in the literature due to its implementation through the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959]. This leaves issues such as the presence of judge' biases hinted by @Bramley_2008 and @Kelly_et_al_2022, and evidenced by @Pollitt_et_al_2003, @Arlett_2003, and @Guthrie_2003 -->

@@

<!-- possibility of bias is hinted in @Bramley_2008, and later mentioned again in @Kelly_et_al_2022, and some evidence is found in @Pollitt_et_al_2003, @Arlett_2003, @Guthrie_2003, @vanDaal_et_al_2017 and @Gijsen_et_al_2021 -->
